---
title: "Exploratory Analysis"
author: "Group 6"
date: "5/17/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
source("analysis.R")
library(ggplot2)
library(dplyr)
library(tidyr)
library(knitr)
```

## Overview
This is a brief report regarding smoking percentage for each states in United Statas from 1995 to 2010. The data reveals that the highest daily smoking percentage in US is Kentucky, with percentage of 24.8% smoking everyday.

Here are the states whose average of daily smoking percentage **are ranked at highest level** from 1995 to 2010. 

```{r tobacco_usage, echo=FALSE}
kable(tobacco, col.names = c("State","Mean (percentage)","Standard Deviation"))
```

## Trend of Top 5 highly-Smoking States Over 15 Years

The following graph presents the trends of smoking from states whose have highest smoking percentages since 1995. The visualization shows that the daily smoking percentage in those top 5 states significantly decrease over the 15 years, with Kentucky having decreased about 6% since 2000. 

```{r trend, echo=FALSE}
ggplot(trend)+
  geom_point(aes(x=Year, y=Smoke.everyday,color=State))
```

